---
title: "Yosemite Recreation Data"
author: "Halina Do-Linh"
format: 
  html:
    embed-resources: true
---

```{r}
#| label: setup
#| echo: false
#| warning: false

library(tidyverse)
library(tidycensus)
library(vroom)
library(here)
library(janitor)
library(zipcodeR)
library(sf)
library(knitr)
library(plotly)
library(scales)
library(tmap)

data_joined_2018 <- readRDS("data-app/2018_joined_data.rds")
## CA prototype joined 2018 metadata ----
data_joined_2018_metadata <- read_csv("data-app/data_joined_2018_metadata.csv")
## CA ACS all ----
data_ca_acs_2018 <- readRDS("data-app/2018_ca_acs_all.rds")
## data relationship 2018 ----
data_education_relationship_plots <- readRDS("data-app/education_relationship_plots.rds")
data_language_relationship_plots <- readRDS("data-app/language_relationship_plots.rds")
data_median_income_relationship_plots <- readRDS("data-app/median_income_relationship_plots.rds")
data_race_relationship_plots <- readRDS("data-app/race_relationship_plots.rds")
## visitorsheds ----
data_state_geometries <- readRDS("data-app/states_geometries.rds")
data_ca_zip_code_geometries <- readRDS("data-app/ca_zip_codes_geometries.rds")
```

In 2018, there were 94,165 overnight campsite reservations made to Yosemite National Park.

```{r}
#| label: all yosemite campsite reservations
#| echo: false

yosemite_all <- data_joined_2018 %>% 
  filter(admin_unit == "Yosemite National Park")
```

See the breakdown of number of reservations made at each Yosemite campground from highest to lowest.

```{r}
#| echo: false
ym_parks <- yosemite_all %>% 
  count(park,
        sort = TRUE) %>% 
  knitr::kable()

ym_parks
```

Since Upper Pines was the most visited campsite at Yosemite in 2018 -- we'll look at the data for that campsite more deeply.

## Booking Window:

```{r}
#| label: booking window function
#| echo: false
not_reactive_booking_window_plot <- function(admin_unit, site, ridb_df){
  
  # data frame
  data <- ridb_df %>%
    filter(park == site,
           booking_window > 0,
           booking_window != "Inf") %>% 
    select(park, booking_window) %>% 
    filter(!is.na(booking_window))
  
  
  # wrangling
  x_max <- numeric(0)
  if(length(data) != 0) {
    
    x_max <- (round(max(data$booking_window) / 5) * 5) + 5 # max x rounded to nearest 5
  }
  
  quant_80 <- numeric()
  if(length(data) != 0) {
    
    quant_80 <- quantile(x = data$booking_window,
                         probs = seq(0, 1, 0.1))[[9]] %>% round(0)
  }
  
  # parameters
  hist_colors <- c("#64863C", "#466C04")
  quant_80_color <- c("#FACE00")
  caption_color <- c("#ac8d00")
  
  plotly <- ggplot(
    data = data) +
    geom_histogram(aes(x = booking_window,
                        text = paste0(scales::percent(after_stat(count) / nrow(data), accuracy = 0.1), 
                                     " of all visits are reserved between ", 
                                     scales::comma(xmin, accuracy = 1), 
                                     " and ", 
                                     scales::comma(xmax, accuracy = 1),
                                     " days before the start of the visit",
                                     "<br>",
                                     "(Total reservations to site: ",
                                     scales::comma(nrow(data), accuracy = 1), ")")),
                   binwidth = 7,
                   center = 7 / 2,
                   fill = hist_colors[[1]], 
                   col = hist_colors[[2]], size = 0.05) +
    labs(x = "Days in advance before visit (each bar = 1 week)",
         y = "") +
    scale_x_continuous(limits = c(0, x_max)) +
    geom_vline(xintercept = quant_80,
               linetype = "dashed", alpha = 0.5, color = quant_80_color) +
    theme_minimal() +
    theme(plot.background = element_rect("white"),
          panel.grid.major.y = element_blank())
  
  # add 6 month / 1 year annotation if data allows
  if (x_max <= 180) {
    # don't include 6 month or 1 year annotation
    plotly
    
  } else if (x_max > 180 & x_max <= 360){
    # include 6 month annotation
    plotly <- plotly +
      geom_vline(xintercept = 180, 
                 linetype = "dashed", linewidth = .3, alpha = .5) +
      annotate("text", label = "6 months",  size = 3,
               x = 180, y = 5)
    
  } else if (x_max >= 360) {
    
    # include 6 month and 1 year annotation
    plotly <- plotly +
      geom_vline(xintercept = 180, 
                 linetype = "dashed", linewidth = .3, alpha = .5) +
      annotate("text", label = "6 months", size = 3,
               x = 180, y = 5) +
      geom_vline(xintercept = 360,
                 linetype = "dashed", linewidth = .3, alpha = .5) +
      annotate("text", label = "1 year", size = 3,
               x = 360, y = 5)
  } # EO else if for plotly
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text"),
           dynamicTicks = TRUE) %>% 
    layout(title = list(text = paste0('<b>', site, '<br>', admin_unit, '</b>',
                                      '<br>',
                                      'Number of days from reservation to start of visit'),
                        font = list(size = 15)),
           xaxis = list(separatethousands = TRUE),
           yaxis = list(separatethousands = TRUE),
           margin = list(b = 130, t = 100), 
           annotations =  list(x = x_max/2, y = -0.6, 
                               text = paste0("80% of reservations reserve their visit less than ", '<b>', quant_80, '</b>', 
                                             " days before the start date", 
                                             "<br>",
                                             "(shown on plot with blue dashed line)."), 
                               showarrow = F, 
                               xre = 'paper', yref = 'paper', 
                               xanchor = 'middle', yanchor = 'auto', 
                               xshift = 0, yshift = 0,
                               font = list(size = 12, color = caption_color))) %>%
    config(modeBarButtonsToRemove = list("pan", "select", "lasso2d", "autoScale2d", 
                                         "hoverClosestCartesian", "hoverCompareCartesian"))
  
  
} # EO function
```


```{r}
#| label: booking window plot
#| echo: false
#| warning: false
not_reactive_booking_window_plot(admin_unit = "Yosemite National Park",
                                 site = "Upper Pines",
                                 ridb_df = data_joined_2018)
```


## Distance Traveled by Median Income

```{r}
#| label: distance traveled function
#| echo: false
not_reactive_median_income_dist_travel_plot <- function(admin_unit,
                                                        site,
                                                        median_income_decile_df) {
  # create dataframe and further subset
  plot_data <- median_income_decile_df %>%
    # filter to user site of choice
    filter(park == site) %>%
    # select to variables of interest
    select(park,
           customer_zip,
           median_income_binned,
           distance_traveled_mi) %>%
    drop_na(median_income_binned) %>%
    # summarize to inner quartile range, median, and total reservations
    group_by(median_income_binned) %>%
    summarize(
      median_distance_traveled_mi = median(distance_traveled_mi),
      quartile_lower = quantile(distance_traveled_mi)[[2]],
      quartile_upper = quantile(distance_traveled_mi)[[4]],
      count = n()
    )
  
  # create plot
  plotly <- ggplot(data = plot_data,
                   aes(x = median_distance_traveled_mi,
                       y = median_income_binned)) +
    geom_segment(aes(xend = 0, yend = median_income_binned)) +
    geom_point(
      aes(
        color = median_income_binned,
        fill = median_income_binned,
        text = paste0(
          scales::comma(count, accuracy = 1),
          " unique visits were made by people who live in ZIP codes with a<br>median household income between ",
          median_income_binned,
          ". Typically these visitors<br>traveled between ",
          scales::comma(quartile_lower, accuracy = 1),
          " and ",
          scales::comma(quartile_upper, accuracy = 1),
          " miles, with a median distance of ",
          scales::comma(median_distance_traveled_mi, accuracy = 1),
          " miles."
        )
      ),
      size = 3.5,
      shape = 21,
      stroke = 2
    ) +
    scale_y_discrete(expand = c(0.2, 0)) +
    scale_fill_viridis_d(direction = -1) +
    scale_color_viridis_d(direction = -1) +
    labs(x = paste("Estimated Distance Traveled from Home to Site (miles)"),
         y = "") +
    theme_minimal() +
    theme(
      plot.background = element_rect("white"),
      panel.grid.major.y = element_blank(),
      legend.position = "none"
    )
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text")) %>%
    config(
      modeBarButtonsToRemove = list(
        "pan",
        "select",
        "lasso2d",
        "autoScale2d",
        "hoverClosestCartesian",
        "hoverCompareCartesian"
      )
    ) %>%
    layout(title = list(
      text = paste0(
        '<b>',
        site,
        '<br>',
        admin_unit,
        '</b>',
        '<br>',
        'Distance Traveled by Visitors with Different Median Household Incomes'
      ),
      font = list(size = 15)
    ))
  
} # EO function
```


```{r}
#| label: distance traveled plot
#| echo: false
#| warning: false
not_reactive_median_income_dist_travel_plot(admin_unit = "Yosemite National Park",
                                            site = "Upper Pines",
                                            median_income_decile_df = data_median_income_relationship_plots)

```

## Median Income

```{r}
#| label: median income function
#| echo: false
not_reactive_median_income_top_quartile_res_plot <- function(admin_unit, 
                                                             site, 
                                                             median_income_decile_df){
  
  # create reactive dataframe and further subset
  plot_data <- median_income_decile_df %>%
      # filter to user site of choice
      filter(park == site) %>%
      # select the variables of interest
      select(park, customer_zip, median_income_binned) %>% 
      drop_na(median_income_binned) %>% 
      # summarize to total reservations
      group_by(median_income_binned) %>% 
      summarize(count = n())
  
  # create plot
  plotly <- ggplot(data = plot_data) +
    geom_col(aes(x = count,
                 y = median_income_binned,
                 fill = median_income_binned,
                 text = paste0(scales::comma(count, accuracy = 1), 
                               " reservations were made<br>by people who live in ZIP codes<br>with median-incomes between<br>", 
                               median_income_binned, "."))) +
    scale_x_continuous(labels = comma_format(accuracy = 1)) +
    scale_y_discrete(expand = c(0.2, 0)) +
    scale_fill_viridis_d(direction = -1) +
    scale_color_viridis_d(direction = -1) +
    labs(x = paste("Number of Reservations"),
         y = "") + 
    theme_minimal() +
    theme(plot.background = element_rect("white"),
          panel.grid.major.y = element_blank(),
          legend.position = "none",
          axis.title.x = element_text(size = 8),
          axis.title.y = element_text(size = 8))
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text"),
           width = 550,
           height = 450) %>%
    config(modeBarButtonsToRemove = list("zoom", "pan", "select", "zoomIn", "zoomOut", "lasso2d", "resetScale2d",
                                         "autoScale2d", "hoverClosestCartesian", "hoverCompareCartesian")) %>% 
    layout(title = list(text = paste0("Reservations to this site from<br>ZIP codes with median-incomes of:"),
                        font = list(size = 14.5)))
  
} # EO function
```

```{r}
#| label: median income plot
#| echo: false
#| warning: false
not_reactive_median_income_top_quartile_res_plot(admin_unit = "Yosemite National Park",
                                            site = "Upper Pines",
                                            median_income_decile_df = data_median_income_relationship_plots)
```

## Distance Traveled by Race

```{r}
#| label: dist travel race function
#| echo: false
not_reactive_race_dist_travel_plot <- function(admin_unit, 
                                               site,
                                               race_top_quartile_df){
  
  # create dataframe and further subset
  plot_data <- race_top_quartile_df %>%
    # filter to user site of choice
    filter(park == site) %>%
    # select to variables of interest
    select(park, customer_zip, 
           race, race_percentage,
           distance_traveled_mi) %>% 
    drop_na(distance_traveled_mi, race_percentage) %>% 
    # summarize to inner quartile range, median, and total reservations
    group_by(race) %>%
    summarize(median_distance_traveled_mi = median(distance_traveled_mi),
              quartile_lower = quantile(distance_traveled_mi)[[2]],
              quartile_upper = quantile(distance_traveled_mi)[[4]],
              count = n())
  
  # parameters
  race_group_colors <- c("Other Race(s)" = "#999999", "Pacific Islander" = "#E69F00", "Multiracial" = "#56B4E9",
                         "Asian" = "#009E73", "Black" = "#F0E442", "White" = "#0072B2", 
                         "Native American" = "#D55E00", "Hispanic Latinx" = "#CC79A7")
  
  # create plot
  plotly <- ggplot(data = plot_data, 
                   aes(x = median_distance_traveled_mi,
                       y = reorder(race, median_distance_traveled_mi))) +
    geom_segment(aes(xend = 0, yend = race)) +
    geom_point(aes(color = race, fill = race,
                   text = paste0(scales::comma(count, accuracy = 1), 
                                 " unique visits were made by people who live in ZIP codes<br>with high ",
                                 race, " populations. Typically these visitors traveled between<br>",
                                 scales::comma(quartile_lower, accuracy = 1), 
                                 " and ", 
                                 scales::comma(quartile_upper, accuracy = 1), 
                                 " miles, with a median distance of ", 
                                 scales::comma(median_distance_traveled_mi, accuracy = 1), 
                                 " miles.")),
               size = 3.5, 
               shape = 21, stroke = 2) +
    scale_y_discrete(expand = c(0.45, 0)) +
    scale_fill_manual(values = race_group_colors) +
    scale_color_manual(values = race_group_colors) +
    labs(x = paste("Estimated Distance Traveled from Home to Site (miles)"),
         y = "") + 
    theme_minimal() +
    theme(plot.background = element_rect("white"),
          panel.grid.major.y = element_blank(),
          legend.position = "none")
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text")) %>%
    config(modeBarButtonsToRemove = list("zoom", "pan", "select", "lasso2d", "autoScale2d", 
                                         "hoverClosestCartesian", "hoverCompareCartesian")) %>% 
    layout(title = list(text = paste0('<b>', site, '<br>', admin_unit, '</b>',
                                      '<br>',
                                      'Distance Traveled by Different Racial Groups'),
                        font = list(size = 15))) %>%
    add_annotations(text = "Reservations from ZIP codes<br>with high proportions of:", 
                    x = -0.5, y = 0.9, 
                    font = list(size = 11),
                    xref = 'paper', yref = 'paper', 
                    showarrow = FALSE)
  
} # EO function
```

```{r}
#| label: dist travel race plot
#| echo: false
#| warning: false
not_reactive_race_dist_travel_plot(admin_unit = "Yosemite National Park",
                                   site = "Upper Pines",
                                   data_race_relationship_plots)
```

## Race

```{r}
#| label: race function
#| echo: false
not_reactive_race_top_quartile_res_plot <- function(admin_unit, 
                                                    site,
                                                    race_top_quartile_df){
  
  # create reactive dataframe and further subset
  plot_data <- race_top_quartile_df %>%
    # filter to user site of choice
    filter(park == site) %>%
    # select to variables of interest
    select(park, customer_zip, 
           race, race_percentage) %>% 
    drop_na(race_percentage) %>% 
    # summarize to inner quartile range, median, and total reservations
    group_by(race) %>%
    summarize(count = n())
  
  # parameters
  race_group_colors <- c("Other Race(s)" = "#999999", "Pacific Islander" = "#E69F00", "Multiracial" = "#56B4E9",
                         "Asian" = "#009E73", "Black" = "#F0E442", "White" = "#0072B2", 
                         "Native American" = "#D55E00", "Hispanic Latinx" = "#CC79A7")
  
  # create plot
  plotly <- ggplot(data = plot_data) +
    geom_col(aes(x = count,
                 y = reorder(race, count),
                 fill = race,
                 text = paste0(scales::comma(count, accuracy = 1), 
                               " reservations were made<br>by people who live in<br>ZIP codes with high<br>", 
                               race, " populations."))) +
    scale_x_continuous(labels = comma_format(accuracy = 1)) +
    scale_y_discrete(expand = c(0.1, 0)) +
    scale_fill_manual(values = race_group_colors) +
    scale_color_manual(values = race_group_colors) +
    labs(x = paste("Number of Reservations"),
         y = "") + 
    theme_minimal() +
    theme(plot.background = element_rect("white"),
          panel.grid.major.y = element_blank(),
          legend.position = "none",
          axis.title.x = element_text(size = 8),
          axis.title.y = element_text(size = 8))
  
  # create plotly
  ggplotly(plotly,
           tooltip = list("text"),
           height = 600,
           width = 450) %>%
    config(modeBarButtonsToRemove = list("zoom", "pan", "select", "zoomIn", "zoomOut", "lasso2d", "resetScale2d",
                                         "autoScale2d", "hoverClosestCartesian", "hoverCompareCartesian")) %>% 
    layout(title = list(text = paste0("Number of Reservations to this site coming from<br>ZIP codes with high proportions of:"),
                        font = list(size = 14.5)))
  
} # EO function
```

```{r}
#| label: race plot
#| echo: false
#| warning: false
not_reactive_race_top_quartile_res_plot(admin_unit = "Yosemite National Park",
                                        site = "Upper Pines",
                                        data_race_relationship_plots)
```

## Visitorshed

```{r}
#| label: national visitorshed function
#| echo: false
not_reactive_state_visitorshed_map <-
  function(site, ridb_df, state_geometries_df) {
    ## -- data wrangle -- ##
    # reactive dataframe of number of reservations per state
    data <- ridb_df %>%
      filter(park %in% site) %>%
      group_by(customer_zip_state_full, customer_zip_state) %>%
      summarize(number_reservations = n()) %>%
      filter(!is.na(customer_zip_state_full))
    
    # add geometries
    map_data <-
      state_geometries_df %>%
      left_join(y = data,
                by = c("state_abbrev" = "customer_zip_state"))
    
    
    ## -- create map -- ##
    tmap_mode("view")
    
    tm_shape(map_data) +
      tm_borders(col = "grey", alpha = 0.5) +
      tm_fill(
        col = "number_reservations",
        title = "Number of Visits",
        palette = "YlGn",
        n = 10,
        style = "jenks",
        id = "customer_zip_state_full",
        popup.vars = c("Total Visits" = "number_reservations")
      ) +
      tm_view(set.view = c(-101.834335, 40.022356, 2.5)) # update zoom
    
  }

```

```{r}
#| label: national visitorshed plot
#| echo: false
#| warning: false
not_reactive_state_visitorshed_map(site = "Upper Pines",
                                   ridb_df = data_joined_2018, 
                                   state_geometries_df = data_state_geometries)
```


















